---
title: "R Notebook of Vancouver Crime"
output: html_notebook
---

## Introduction
This is the instruction for the codes of UCL MSc [CEGE0042](https://moodle.ucl.ac.uk/course/view.php?id=1329&section=0#tabs-tree-start) coursework. The aim of this coursework is to conduct an exploratory spatio-temporal analysis of Vancouver crime data and make predictions.  
  
Most of the code is derived from Dr James' book [Spatio-temporal Analytics in R](https://moodle.ucl.ac.uk/pluginfile.php/1301215/mod_resource/content/17/_book/index.html).  
Data source: [Crime in Vancouver](https://www.kaggle.com/datasets/wosaku/crime-in-vancouver), [Vancouver district](https://opendata.vancouver.ca/explore/dataset/local-area-boundary/export/?disjunctive.name). 
  
Directories in the coursework will point to a folder called 'Coursework'. Please create this folder somewhere in your computer and use 'setwd' to set the working directory to point to it. For example, if 'Coursework' is placed in '/Users' then run setwd('/Users/Coursework').  
  
**This notebook is divided into 3 parts:**  
- Part 1: Data Import and Pre-processing  
- Part 2: Data Visualization & Autocorrelation  
- Part 3: ARIMA & ANN  

## Part 1: Data Import and Pre-processing
First check and change working directory using *getwd()* & *setwd()*. Import Vancouver Crime data and select interested attributes. Using *head()* to see the structure of the data. Then select the data we gonna to deal with.
```{r}
# check and change working directory using getwd() & setwd()
getwd()
setwd("/Users/zmxu/Desktop/UCL MSc Geospatial Sciences/term2-周2-CEGE0042-Spatial-Temporal Data Analysis and Data Mining (STDM) /Coursework")

# import Vancouver Crime data
data <- read.csv("archive/crime.csv")
head(data,10)

# select interested attributes
crime <- data.frame(data[,c("TYPE","YEAR","MONTH","NEIGHBOURHOOD","Latitude","Longitude")])
```
There might be some recording errors in the data set. Clean the data set in case there are NA & 0, and output the first several lines.
```{r}
crime[crime==0] <- NA
crime <- na.omit(crime)
print(nrow(crime))
head(crime,10)
```
In order to read crime data by different type and neighbourhood, a zero-value matrix is created to hold the data, and the number of different types of crime is counted and written to the matrix by cyclically filtering the crime events that meet the conditions.  
*Actually the TYPE of crime was also selected, in order to know more about the data and would be helpful to the further work.*
```{r}
# check different TYEP & NEIGHBOURHOOD
unique(crime$TYPE)
unique(crime$NEIGHBOURHOOD)

# create year_month matrix (might introduce 0)
# all crimes
year_month <- matrix(data=0,nrow=15,ncol=12)
# group by crime TYPE
year_month_M     <- matrix(data=0,nrow=15,ncol=12)
year_month_OT    <- matrix(data=0,nrow=15,ncol=12)
year_month_ToV   <- matrix(data=0,nrow=15,ncol=12)
year_month_TfV   <- matrix(data=0,nrow=15,ncol=12)
year_month_ToB   <- matrix(data=0,nrow=15,ncol=12)
year_month_BEC   <- matrix(data=0,nrow=15,ncol=12)
year_month_BERO  <- matrix(data=0,nrow=15,ncol=12)
year_month_VCPSI <- matrix(data=0,nrow=15,ncol=12)
year_month_VCPSF <- matrix(data=0,nrow=15,ncol=12)
# group by NEIGHBOURHOOD
year_month_Strathcona              <- matrix(data=0,nrow=15,ncol=12)
year_month_Kerrisdale              <- matrix(data=0,nrow=15,ncol=12)
year_month_DunbarSouthlands        <- matrix(data=0,nrow=15,ncol=12)
year_month_GrandviewWoodland       <- matrix(data=0,nrow=15,ncol=12)
year_month_Sunset                  <- matrix(data=0,nrow=15,ncol=12)
year_month_WestEnd                 <- matrix(data=0,nrow=15,ncol=12)
year_month_CentralBusinessDistrict <- matrix(data=0,nrow=15,ncol=12)
year_month_HastingsSunrise         <- matrix(data=0,nrow=15,ncol=12)
year_month_VictoriaFraserview      <- matrix(data=0,nrow=15,ncol=12)
year_month_Fairview                <- matrix(data=0,nrow=15,ncol=12)
year_month_KensingtonCedarCottage  <- matrix(data=0,nrow=15,ncol=12)
year_month_WestPointGrey           <- matrix(data=0,nrow=15,ncol=12)
year_month_Shaughnessy             <- matrix(data=0,nrow=15,ncol=12)
year_month_RenfrewCollingwood      <- matrix(data=0,nrow=15,ncol=12)
year_month_Killarney               <- matrix(data=0,nrow=15,ncol=12)
year_month_RileyPark               <- matrix(data=0,nrow=15,ncol=12)
year_month_ArbutusRidge            <- matrix(data=0,nrow=15,ncol=12)
year_month_Musqueam                <- matrix(data=0,nrow=15,ncol=12)
year_month_MountPleasant           <- matrix(data=0,nrow=15,ncol=12)
year_month_Kitsilano               <- matrix(data=0,nrow=15,ncol=12)
year_month_StanleyPark             <- matrix(data=0,nrow=15,ncol=12)
year_month_SouthCambie             <- matrix(data=0,nrow=15,ncol=12)
year_month_Marpole                 <- matrix(data=0,nrow=15,ncol=12)
year_month_Oakridge                <- matrix(data=0,nrow=15,ncol=12)

year  <- c(2003:2017)
month <- c(1:12)

for (i in year){
  for (j in month){
    # all crimes
    year_month[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j))
    # group by TYPE
    year_month_M[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Mischief"))
    year_month_OT[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Other Theft"))
    year_month_ToV[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Theft of Vehicle"))
    year_month_TfV[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Theft from Vehicle"))
    year_month_ToB[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Theft of Bicycle"))
    year_month_BEC[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Break and Enter Commercial"))
    year_month_BERO[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Break and Enter Residential/Other"))
    year_month_VCPSI[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Vehicle Collision or Pedestrian Struck (with Injury)"))
    year_month_VCPSF[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & TYPE=="Vehicle Collision or Pedestrian Struck (with Fatality)"))
    # group by NEIGHBOURHOOD
    year_month_Strathcona[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Strathcona"))
    year_month_Kerrisdale[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Kerrisdale"))
    year_month_DunbarSouthlands[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Dunbar-Southlands"))
    year_month_GrandviewWoodland[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Grandview-Woodland"))
    year_month_Sunset[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Sunset"))
    year_month_WestEnd[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="West End"))
    year_month_CentralBusinessDistrict[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Central Business District"))
    year_month_HastingsSunrise[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Hastings-Sunrise"))
    year_month_VictoriaFraserview[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Victoria-Fraserview"))
    year_month_Fairview[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Fairview"))
    year_month_KensingtonCedarCottage[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Kensington-Cedar Cottage"))
    year_month_WestPointGrey[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="West Point Grey"))
    year_month_Shaughnessy[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Shaughnessy"))
    year_month_RenfrewCollingwood[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Renfrew-Collingwood"))
    year_month_Killarney[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Killarney"))
    year_month_RileyPark[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Riley Park"))
    year_month_ArbutusRidge[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Arbutus Ridge"))
    year_month_Musqueam[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Musqueam"))
    year_month_MountPleasant[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Mount Pleasant"))
    year_month_Kitsilano[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Kitsilano"))
    year_month_StanleyPark[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Stanley Park"))
    year_month_SouthCambie[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="South Cambie"))
    year_month_Marpole[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Marpole"))
    year_month_Oakridge[i-2002,j] <- nrow(subset(crime, YEAR==i & MONTH==j & NEIGHBOURHOOD=="Oakridge"))
  }
}

# all Crime - change row&col name
colnames(year_month)[1:ncol(year_month)] <- as.character(c(1:12))
rownames(year_month)[1:nrow(year_month)] <- as.character(c(2003:2017))
```
The matrix is transformed into a time series and intercepted up to the first 175 items for which there are actual observations.
```{r}
# covert year_month matrix to month series (delete 0)
# all crime
time_series_long <- as.vector(t(year_month))
time_series      <- time_series_long[1:175]
# group by TYPE
time_series_long  <- as.vector(t(year_month_M)) # TYPE=="Mischief"
time_series_M     <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_OT)) # TYPE=="Other Theft"
time_series_OT    <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_ToV)) # TYPE=="Theft of Vehicle"
time_series_ToV   <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_TfV)) # TYPE=="Theft from Vehicle"
time_series_TfV   <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_ToB)) # TYPE=="Theft of Bicycle"
time_series_ToB   <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_BEC)) # TYPE=="Break and Enter Commercial"
time_series_BEC   <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_BERO)) # TYPE=="Break and Enter Residential/Other"
time_series_BERO  <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_VCPSI)) # TYPE=="Vehicle Collision or Pedestrian Struck (with Injury)"
time_series_VCPSI <- time_series_long[1:175]
time_series_long  <- as.vector(t(year_month_VCPSF)) # TYPE=="Vehicle Collision or Pedestrian Struck (with Fatality)"
time_series_VCPSF <- time_series_long[1:175]
# group by NEIGHBOURHOOD
time_series_long <- as.vector(t(year_month_Strathcona))
time_series_Strathcona <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Kerrisdale))
time_series_Kerrisdale <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_DunbarSouthlands))
time_series_DunbarSouthlands <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_GrandviewWoodland))
time_series_GrandviewWoodland <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Sunset))
time_series_Sunset <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_WestEnd))
time_series_WestEnd <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_CentralBusinessDistrict))
time_series_CentralBusinessDistrict <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_HastingsSunrise))
time_series_HastingsSunrise <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_VictoriaFraserview))
time_series_VictoriaFraserview <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Fairview))
time_series_Fairview <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_KensingtonCedarCottage))
time_series_KensingtonCedarCottage <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_WestPointGrey))
time_series_WestPointGrey <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Shaughnessy))
time_series_Shaughnessy <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_RenfrewCollingwood))
time_series_RenfrewCollingwood <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Killarney))
time_series_Killarney <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_RileyPark))
time_series_RileyPark <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_ArbutusRidge))
time_series_ArbutusRidge <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Musqueam))
time_series_Musqueam <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_MountPleasant))
time_series_MountPleasant <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Kitsilano))
time_series_Kitsilano <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_StanleyPark))
time_series_StanleyPark <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_SouthCambie))
time_series_SouthCambie <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Marpole))
time_series_Marpole <- time_series_long[1:175]
time_series_long <- as.vector(t(year_month_Oakridge))
time_series_Oakridge <- time_series_long[1:175]
```
Combine the data of different neighbourhoods into a matrix, rewrite rownames according to the neighbourhood.shp file,rewrite colnames by month, and output the result as a .csv file.
```{r}
# combine all district data to one matrix
time_series_by_neighbourhood <- rbind(time_series_Strathcona,time_series_Kerrisdale,time_series_DunbarSouthlands,
                                      time_series_GrandviewWoodland,time_series_Sunset,time_series_WestEnd,
                                      time_series_CentralBusinessDistrict,time_series_HastingsSunrise,time_series_VictoriaFraserview,
                                      time_series_Fairview,time_series_KensingtonCedarCottage,time_series_WestPointGrey,
                                      time_series_Shaughnessy,time_series_RenfrewCollingwood,time_series_Killarney,
                                      time_series_RileyPark,time_series_ArbutusRidge,time_series_Musqueam,
                                      time_series_MountPleasant,time_series_Kitsilano,time_series_StanleyPark,
                                      time_series_SouthCambie,time_series_Marpole,time_series_Oakridge)
# rename district name according to .shp file
rownames(time_series_by_neighbourhood)[1:nrow(time_series_by_neighbourhood)] <- as.character(c("Strathcona","Kerrisdale","Dunbar-Southlands",
                                                                                               "Grandview-Woodland","Sunset","West End",
                                                                                               "Downtown","Hastings-Sunrise","Victoria-Fraserview",
                                                                                               "Fairview","Kensington-Cedar Cottage","West Point Grey",
                                                                                               "Shaughnessy","Renfrew-Collingwood","Killarney",
                                                                                               "Riley Park","Arbutus-Ridge",  "Musqueam",
                                                                                               "Mount Pleasant","Kitsilano",  "StanleyPark",
                                                                                               "South Cambie","Marpole","Oakridge"))
# set month number
Month.number <- paste("Month ",1:175,sep="")
colnames(time_series_by_neighbourhood) <- as.character(Month.number)

# output to .csv, in order to link in ArcGIS
write.csv(x = time_series_by_neighbourhood,file = "time_series_by_neighbourhood.csv")
```
There are 2 steps needs to be done before continue the codes:  
1. Link the rime data in *"time_series_by_neighbourhood.csv"* to neighborhood data (.shp) by ArcGIS software.  
2. Use the output to substitute the *"local_area_boundary.dbf"* file in the *"Vancouver-local-area-boundary"* folder.  
After that, continue the following codes.  

*Actually, "local_area_boundary.dbf" file in the "Vancouver-local-area-boundary" folder has already linked by the author in advance, so you can continue the work now.*


## Part 2: Data Visualization & Autocorrelation
### Data Visualization - Non Spatio-temporal data characteristics
After pre-processing operations such as loading, arranging and cleaning the data, the temporal distribution of crime incidents in Vancouver from 2003.1.1 to 2017.7.13 could be plotted and basic mathematical statistical information could be presented.  
This part of code will produce *Figure 2.1 Vancouver Crime Time Series and Statistical Information*
```{r,out.width='100%',out.height='100%',fig.height=3, fig.width=9, fig.align="left"}
par(mfrow=c(1,3))
# Time series plot
plot(time_series, xlab = "Year", ylab = "Crime Number", type="l", xaxt="n", main="Time series of Crime by Month")
axis(1, at=seq(1,180,12), labels=seq(2003,2017,1))
# Histogram
mu = mean(time_series)
hist(time_series, main="Histogram of Crime (Monthly)")
abline(v=mu, col="red")
# Q-Q plot
qqnorm(time_series, main="Normal Q-Q Plot of Crime (Monthly)")
qqline(time_series, col="red")
```
### Data Visualization - Spatial characteristics
First load some packages. Install if necessary.
```{r}
library(ggplot2)
library(OpenStreetMap)
library(raster)
library(ggmap)
library(osmdata)
```
Crime in the dataset is recorded as a scatter and the 2003.10 data has been selected for presentation. In addition, it is possible to obtain the crime density of the different subdivisions.  
This part of code will produce *Figure 1.1 Spatial Distribution of Crime in Vancouver (2003.10)*
```{r,out.width='100%',out.height='100%',fig.height=3, fig.width=3, fig.align="left"}
par(mfrow=c(1,1))
# Figure: Crime in Vancouver, 2003.10
# add crime data: eg.2016 data
crime_data <- data.frame(subset(crime, YEAR==2003 & MONTH==10))
# plot Vancouver map
mad_map <- get_map(getbb("Vancouver"), maptype = "toner-background")
ggmap(mad_map)+
  geom_point (data=crime_data, aes(x=Longitude,y=Latitude), color = "#BA4A00",size=0.3)+
  ggtitle("Crime in Vancouver, 2003.10")+
  labs(x = "LON", y = "LAT")

# Figure: Crime distribution in 2003.10
library(rgdal)
library(tmap)
vancouver_districts <- readOGR(dsn="Vancouver-local-area-boundary/local_area_boundary.shp") # local_area_boundary.shp has already linked crime data
#tmap_style("cobalt") # set style
#tmap_style("white") # back to original
tm_shape(vancouver_districts)+ 
  tm_fill("Month_10", style="jenks", palette="Blues")+
  tm_borders("white")+
  tm_compass(position=c("left","top"))+
  tm_scale_bar(position=c(0.6,0.005))
```
### Temporal autocorrelation
More in-depth exploration of crime data was conducted to calculate temporal autocorrelations for different time scales and different sub-districts. For the time scales, the time series of Vancouver in terms of months and years were selected to calculate ACF and PACF. Two neighbourhoods, Downtown and Marpole, where the number of crime incidents differed significantly, were selected for comparison.  
This part of code will produce *Figure 2.2 Temporal autocorrelation: ACF&PACF of different time and space scales*
```{r,out.width='100%',out.height='100%',fig.height=5, fig.width=10, fig.align="left"}
Downtown <- time_series_by_neighbourhood[7,1:175]
Marpole <- time_series_by_neighbourhood[23,1:175]
par(mfrow=c(2,4))
acf(colMeans(matrix(time_series,12)), main="ACF, Annual")
acf(time_series,lag.max=35, main="ACF, Monthly")
acf(Downtown,lag.max=35, main="ACF, Monthly, Downtown")
acf(Marpole,lag.max=35, main="ACF, Monthly, Marpole")
pacf(colMeans(matrix(time_series,12)), main="PACF, Annual")
pacf(time_series,lag.max=35, main="PACF, Monthly")
pacf(Downtown,lag.max=35, main="PACF, Monthly, Downtown")
pacf(Marpole,lag.max=35, main="PACF, Monthly, Marpole")
```
### Spatial autocorrelation
Use Moran's I to evaluate Global Spatial Autocorrelation.
```{r}
library(spdep)
W <- nb2listw(poly2nb(vancouver_districts))
W
library(knitr)
kable(listw2mat(W))

# Global Spatial Autocorrelation Measures
# Moran Index
crime_matrix<-data.matrix(vancouver_districts@data[,-c(1:3)])
rownames(crime_matrix) <- vancouver_districts@data[,"name"]
crime_avg <- rowMeans(crime_matrix)
moran.test(x=crime_avg, listw=W) # Moran I 0.25908 low
moran.mc(x=crime_avg, listw=W, nsim=9999) # another way to calculate Moran I
```
Spatial autocorrelation can be demonstrated by calculating and plotting Moran scatter plots.  
This part of code will produce *Figure 2.3 Spatial autocorrelation: Moran scatter plot & LISA plot*
```{r,out.width='100%',out.height='100%',fig.height=5, fig.width=5, fig.align="left"}
# Figure: Moran scatter plot
par(mfrow=c(1,1))
moran.plot(crime_avg,W,main="Morans'I = 0.25908")

# local Moran's I
lm <- localmoran(x=rowMeans(crime_matrix), listw=W)
lm
# Figure: LISA （this part is done by GeoDa）
```
### Spatio-temporal autocorrelation
In order to demonstrate the degree of Spatio-temporal autocorrelation, a hotspot map was plotted, and Spatio-temporal Autocorrelation and Spatio-temporal Partial Autocorrelation were also calculated.  
This part of code will produce *Figure 2.4 Spatio-temporal autocorrelation*
```{r,out.width='100%',out.height='100%',fig.height=3, fig.width=6, fig.align="left"}
# Figure: Heatmap
library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "Blues"))(25)
heatmap(crime_matrix,Colv = NA, Rowv = NA, scale="column", col = coul, xlab="", ylab="", main="Heatmap")

source("starima_package.R")
Wmat <- listw2mat(W)
# Figure: STACF & STPACF
par(mfrow=c(1,2))
stacf(t(crime_matrix), Wmat, 48)
stpacf(t(crime_matrix), Wmat, 4)
```

## Part 3: ARIMA & ANN
In this coursework, I tried to use ARIMA and ANN model to predict the Crime in Vancouver.
### ARIMA Model
The Autoregressive Integrated Moving Average (ARIMA) model, is a class of models that interprets a given time series based on the historical values of the data. Any non-seasonal time series with a certain pattern that is not random white noise can be modelled and predicted by ARIMA.  
ARIMA modelling according to the Box-Jenkins method has the following main steps:  
a. Exploratory data analysis  
b. Autocorrelation and partial autocorrelation analysis  
c. Model Identification  
d. Parameter Estimation and Fitting  
e. Diagnostic Checking  
f. Forecasting/Prediction  
The data was first divided into a training set (80%) and a test set (20%). Based on the training set, exploratory data analysis was carried out and ACF, as well as PACF, were calculated.
```{r}
library(TSA)
library(forecast)

# train set & test set
time_series_train <- time_series[1:140]
time_series_test <- time_series[141:175]
```
Then try to plot ACF and PACF to show the temporal autocorrelation.  
This part of code will produce *Figure 3.1 Train set temporal autocorrelation(before differencing)*
```{r,out.width='100%',out.height='100%',fig.height=5, fig.width=5, fig.align="left"}
par(mfrow=c(1,1))
# Figure: time_series & ACF & PACF
tsdisplay(time_series_train,xlab="Time (month)",ylab="Crime Number")
# Figure: lag 1 to 12
lag.plot(time_series_train, lags=12, do.lines=FALSE)
```
This part of code will produce *Figure 3.2 Train set temporal autocorrelation(after differencing)*
```{r,out.width='100%',out.height='100%',fig.height=3, fig.width=3, fig.align="left"}
# Figure: diff=1, time_series & ACF & PACF
time_series_diff1 <- diff(time_series_train, lag=1, differences=1)
tsdisplay(time_series_diff1,xlab="Time (month)",ylab="Crime Number")
```
Estimate parameters by the figures above that ARIMA(1,1,1) might be suitable. Output the summary of this model. The parameters can also be automatically discriminated by using the *auto.arima()* function of the R Forecast package.
```{r}
# estimate by figure 12 & 14: ARIMA(1,1,1)
fit.arima111 <- Arima(time_series_train,order=c(1,1,1))
summary(fit.arima111)

# determine parameters automatically
fit.arima <- auto.arima(time_series_train,trace=T)
fit.arima # ARIMA(1,1,1)

summary(fit.arima) # model summary
```
Based on the ARIMA (1, 1, 1) model future data can be predicted and compared with the test set.  
This part of code will produce *Figure 3.4 Vancouver Crime Prediction Results by ARIMA*
```{r,out.width='100%',out.height='100%',fig.height=3, fig.width=6, fig.align="left"}
# Figure: forcast & contrast
fit.Ar <- Arima(time_series_train, order=c(1,1,1))
pre.Ar <- Arima(time_series_test, model=fit.Ar)
matplot(cbind(pre.Ar$x, pre.Ar$fitted),ylab="Crime Number", xlab="Month", main="Vancouver Crime, ARIMA",type="l")
legend("bottomleft", title="Data", c("Real","Predict"), lty=c(1, 2), col=c("black", "red"))
```
This part of code will produce *Figure 3.2 Residual distribution*
```{r,out.width='100%',out.height='100%',fig.height=5, fig.width=5, fig.align="left"}
# Figure: Model Residuals
checkresiduals(fit.arima) # residuals 1
tsdiag(fit.arima) # residuals 2
```
Calculate the NRMSE by function *NRMSE()* of "starima_package.R"
```{r}
source("starima_package.R")
NRMSE_ARIMA <- NRMSE(res=fit.Ar$residuals, obs=time_series_train)
NRMSE_ARIMA

NRMSE_ANN <- NRMSE(obs=time_series_test, pred=pre.Ar$fitted)
NRMSE_ARIMA
```

### ANN Model
The Artificial Neural Network (ANN) is a simulation of biological neurons, with a model that resembles the human brain and consists of a large number of processors, which is more effective in predicting spatial and temporal changes in crime and disorder events. The same method was used to divide the data into a train set and a test set, and a multiple input, multiple output ANN model was constructed, using all the data from the previous month to predict all the data for the following month.
Load the necessary packages. Same method was used to divide the data into a train set and a test set.
```{r}
library(nnet)
library(rgdal)
vancouver_districts <- readOGR(dsn="Vancouver-local-area-boundary/local_area_boundary.shp")
crime_matrix<-data.matrix(vancouver_districts@data[,-c(1:3)])
rownames(crime_matrix) <- vancouver_districts@data[,"name"]
X <- t(as.matrix(crime_matrix))
y <- as.matrix(X[-1,])
```
In order to compare the fitting results for different neighbourhoods of the same method, forecasts were made for Downtown and Marpole using the ANN model.  
This part of code will produce *Figure 4.1 Prediction results of different regions and models*
```{r}
par(mfrow=c(2,2))
# Figure: ANN plot, Downtown
set.seed(319)
crime.nnet <- nnet(X[1:140, 1:22], y[1:140, 1:22], decay=0.01, linout = TRUE, size=6, maxit = 1000)
crime.pred <- predict(crime.nnet, y[141:174, 1:22])
crime.pred[,8]
matplot(cbind(y[141:174,8], crime.pred[,8]),ylab="Crime Number", xlab="Month", main="Downtown Crime, ANN", type="l")

NRMSE_ANN <- NRMSE(obs=X[141:174,8],pred=crime.pred[,8])
NRMSE_ANN

# Figure: ANN plot, Marpole
set.seed(319)
crime.nnet <- nnet(X[1:140, 1:22], y[1:140, 1:22], decay=0.01, linout = TRUE, size=6, maxit = 1000)
crime.pred <- predict(crime.nnet, y[141:174, 1:22])
crime.pred[,12]
matplot(cbind(y[141:174,12], crime.pred[,12]),ylab="Crime Number", xlab="Month", main="Marpole Crime, ANN", type="l")
NRMSE_ANN <- NRMSE(obs=X[141:174,12],pred=crime.pred[,12])
NRMSE_ANN

# create Vancouver Total Crime numbers of each month
X.sum <- as.matrix(rowSums(X))
X.sum <- cbind(X, X.sum)
colnames(X.sum)[ncol(X.sum)] <- as.character(c("Vancouver"))
y.sum <- as.matrix(X.sum[-1,])

# Figure: ANN plot, Vancouver
set.seed(319)
crime.nnet <- nnet(X.sum[1:140, 1:23], y.sum[1:140, 1:23], decay=10, linout = TRUE, size=6, maxit = 100)
crime.pred <- predict(crime.nnet, y.sum[141:174, 1:23])
crime.pred[,23]
matplot(cbind(y.sum[141:174,23], crime.pred[,23]),ylab="Crime Number", xlab="Month", main="Vancouver Crime, ANN", type="l")
# legend("bottomleft", title="Data", c("Real","Predict"), lty=c(1, 2), col=c("black", "red"))

NRMSE_ANN <- NRMSE(res=crime.nnet$residuals, obs=X.sum[1:140, 1:23])
NRMSE_ANN

NRMSE_ANN <- NRMSE(obs=X.sum[141:174,23],pred=crime.pred[,23])
NRMSE_ANN

matplot(cbind(pre.Ar$x, pre.Ar$fitted),ylab="Crime Number", xlab="Month", main="Vancouver Crime, ARIMA",type="l")
```

This part of code will produce *Figure 3.5 Internal structure of ANN*
```{r,out.width='100%',out.height='100%',fig.height=5, fig.width=14, fig.align="left"}
par(mfrow=c(1,1))
# Figure: Visualisation of internal structure of ANN
library(devtools)
source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff044412703516c34f1a4684a5/nnet_plot_update.r')
plot.nnet(crime.nnet)
summary(crime.nnet) # check the weights
```

This part of code will produce *Figure 3.6 Predictions vs Observed*
```{r,out.width='100%',out.height='100%',fig.height=5, fig.width=5, fig.align="left"}
# Figure: Scatterplot of observed vs predicted values
plot(crime.pred,y.sum[141:174, 1:23], main="ANN predictions vs observed, Vancouver", xlab="Observed", ylab="Predicted")
```
























